\sysfull

### 2018-08-03 paper/asplos19/background.tex
New WAR figure (and explanation)

% \begin{wrapfigure}{t!}{0.5\textwidth}
\begin{figure}
	\begin{subfigure}[t]{\linewidth}
		\centering \includegraphics[width=\columnwidth]{figures/crc_example}
		\caption{Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration.}\label{fig:crc_example}
	\end{subfigure}
	\begin{subfigure}[t]{\linewidth}
		\centering \includegraphics[width=\columnwidth]{figures/crc_example_war}
		\caption{Execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)}\label{fig:crc_example_war}
	\end{subfigure}
	\caption{Code example demonstrating checkpointing gives rise to data inconsistency.}\label{fig:code_demo_incosistency}
% \end{wrapfigure}
\end{figure}

@@

Checkpointing volatile state alone do not ensure data consistency when the
system can directly manipulate non-volatile memory~\cite{mspcdino}.
Especially, data can get inconsistent when code includes a
\emph{write-after-read} (WAR) dependence between operations that manipulate
non-volatile memory, because it allows {\em writes} from a failed attempt to
persist across failure and be read by the {\em
reads}~\cite{ratchet,dino,alpaca}\TODO{figure 2 dependent text ---see
Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become
inconsistent in an intermittent execution using the cyclic redundancy check
(CRC) code from MIBench2~\cite{hicks_mibench2_2016}.  Additional method to
maintain data consistent in the presence of WAR dependency is necessary for
correct execution.}
%Checkpoints of volatile state preserve intermittent progress, but do not ensure data consistency~\cite{dino,chain,ratchet}. Data may become inconsistent if an attempt to execute some computation {\em writes} to a non-volatile variable, then power fails, then a second attempt to re-execute the same computation incorrectly {\em reads} the value written in the first attempt, rather than the variable's original value. The situation occurs when code includes a \emph{write-after-read} (WAR) dependence between operations that manipulate non-volatile variables~\cite{ratchet,dino,alpaca}---see Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
%Figure~\ref{fig:code_demo_incosistency} illgit statuustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}. The code computes the CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
%Despite checkpointing at line 3 in Fig.~\ref{fig:crc_example}, the code may compute  $\texttt{data}$ incorrectly because of the read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5. Figure~\ref{fig:crc_example_war} shows an intermittent execution. The second iteration writes $\texttt{Rmnd}$, but power fails before the checkpoint. After restarting, line 9 reads the updated value of $\texttt{Rmnd}$, producing an incorrect $\texttt{data}$ value. Checkpoint-based systems risk violating memory consistency~\cite{dino}. 
%To maintain consistency, prior systems~\cite{dino} {\em version} a subset of non-volatile data with the checkpoint. Task-based systems~\cite{chain,alpaca}, which we focus on, ensure consistency using programmer, compiler, and runtime support.

>>>>

\begin{figure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
        \caption{WAR-affected code}
        \label{fig:war-example}
    \end{subfigure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
        \caption{Incorrect execution}
        \label{fig:war-execution}
    \end{subfigure}
    \caption{WAR dependence example. \texttt{NV} marks non-volatile variables,
    while \texttt{checkp()} is a checkpoint of volatile state.}
    \label{fig:war}
\end{figure}

@@

Checkpointing volatile state alone does not ensure data consistency when the
system can directly manipulate non-volatile memory~\cite{mspcdino}.
Precisely, data can get inconsistent when code includes a
\emph{write-after-read} (WAR) dependence between operations that manipulate
non-volatile memory, because it allows {\em writes} from a failed attempt to
persist across failure and be read by the {\em
reads}~\cite{ratchet,dino,alpaca}.
Figure~\ref{fig:war} illustrates how state can become
inconsistent in an intermittent execution using a simple example of an average
operation over an array of integers.
The variable \texttt{sum} introduces the WAR, being read and written sequentially
by the increment operation, thus being erroneously incremented when a power
failure occurs before updating the index \texttt{i}.
